### 2022-02-14 chapters/conclusions/conclusions.tex
encore plus de refs

Our coupled hydro-epidemiological model reproduced the observed propagation of the SCTLD through the FRT by representing the transport of its causative agent within neutrally buoyant material driven by mean barotropic currents. It confirmed the result of previous ex-situ studies that showed evidence of waterborne transmission of the SCTLD with an average transmission time of the order of 10 days. Furthermore, by using coral resistance to the disease as a parameter, our model results suggested that, on average, corals had a rather low resistance against the disease. This confirmation of experimental results by model results illustrates the important potential of models in evaluating hypotheses derived from field observations on large scale systems. As such, models informed by and confronted against field knowledge are a powerful tool to inform the management of complex ecosystems such as the FCR.

@@

\subsection*{Combining disease and larval connectivity}
All chapters dedicated to SCTLD rely on the concept of connectivity. In this context, connectivity negatively impacts coral reefs as it drives the propagation of disease agents. However, connectivity is actually a double-edged sword. Larval connectivity promotes the resilience of the reef network as it allows the recolonization of disturbed reefs following extreme events such as mass bleaching, hurricanes or disease outbreaks. Larval connectivity is therefore a valuable tool to inform coral reef protection and restoration. Identifying reefs that are both good sources and sinks of larvae the connectivity network indicates the optimal sites for coral outplants in the context of active restoration. Furthermore, knowledge about larval connectivity would allow us to build robust and sustainable networks of connected marine protected areas. As SLIM allows the computation of larval connectivity \citep{thomas2014numerical,frys2020fine}, a possible application of the tools developed in this thesis is the combination of both types of connectivity.  This combined information would allow us to identify reefs that are good larval sources and have low vulnerability to the disease. A possible measure of larval supply to the network is the (weighted) out-degree of reefs in the larval connectivity network, as reefs with strong and numerous outgoing connections in the network are more likely to send many larvae to many connected reefs. Vulnerability to disease can be evaluated using the (weighted) in-degree of reefs in the disease connectivity network, as reefs with strong and numerous incoming connections are more likely to receive more disease agents from many connected reefs. Therefore, reefs with high restoration potential would correspond to reefs with large larval(weighted) out-degree and low disease (weighted) in-degree. This approach have been applied in \cite{holstein2022} and could be further developed in future studies using more complex connectivity metrics such as the PageRank index \citep{frys2020fine}.

>>>>

Our coupled hydro-epidemiological model reproduced the observed propagation of the SCTLD through the FRT by representing the transport of its causative agent within neutrally buoyant material driven by mean barotropic currents. It confirmed the result of previous ex-situ studies that showed evidence of waterborne transmission of the SCTLD with an average transmission time of the order of 10 days. Furthermore, by using coral resistance to the disease as a parameter, our model results suggested that, on average, corals had a rather low resistance against the disease. This confirmation of experimental results by model results illustrates the important potential of models in evaluating hypotheses derived from field observations on large scale systems. As such, models informed by and confronted against field knowledge are a powerful tool to inform the management of complex ecosystems such as the FCR \citep{foster2012connectivity}.

@@

\subsection*{Combining disease and larval connectivity}
All chapters dedicated to SCTLD rely on the concept of connectivity. In this context, connectivity negatively impacts coral reefs as it drives the propagation of disease agents. However, connectivity is actually a double-edged sword. Larval connectivity promotes the resilience of the reef network as it allows the recolonization of disturbed reefs following extreme events such as mass bleaching, hurricanes or disease outbreaks. Larval connectivity is therefore a valuable tool to inform coral reef protection and restoration \citep{botsford2009connectivity,mumby2011reserve}. Identifying reefs that are both good sources and sinks of larvae the connectivity network indicates the optimal sites for coral outplants in the context of active restoration. Furthermore, knowledge about larval connectivity would allow us to build robust and sustainable networks of connected marine protected areas. As SLIM allows the computation of larval connectivity \citep{thomas2014numerical,frys2020fine}, a possible application of the tools developed in this thesis is the combination of both types of connectivity.  This combined information would allow us to identify reefs that are good larval sources and have low vulnerability to the disease. A possible measure of larval supply to the network is the (weighted) out-degree of reefs in the larval connectivity network, as reefs with strong and numerous outgoing connections in the network are more likely to send many larvae to many connected reefs. Vulnerability to disease can be evaluated using the (weighted) in-degree of reefs in the disease connectivity network, as reefs with strong and numerous incoming connections are more likely to receive more disease agents from many connected reefs. Therefore, reefs with high restoration potential would correspond to reefs with large larval(weighted) out-degree and low disease (weighted) in-degree. This approach have been applied in \cite{holstein2022} and could be further developed in future studies using more complex connectivity metrics such as the PageRank index \citep{frys2020fine}.
